---
title: 'dplyr: basic functions'
author: "Samuel Macêdo"
date: '2018-04-26'
categories: ["tutorials"]
tags: ["dplyr"]
description: "Take your first step in data manipulation with dplyr package. This tutorial will present you the basic functions: select(), filter(), mutate(), transmute(), group_by() and summarise()."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Welcome, friend :)

In this tutorial, I am going to present you the basic functions in the dplyr package: select(), filter(), mutate(), transmute(), group_by() and summarise().

If you have any doubts, don't hesitate to contact me on `samuelmacedo@recife.ifpe.edu.br`. 

Let's get to action...first of all To install it, please use the commands below: 

```{r, eval = FALSE}
install.packages("dplyr")
library(dplyr)
```

```{r include=FALSE}
library(dplyr)
```

### Before I start: as_tibble()

Tibble is a modern reimagining of the data.frame. You don't need to change your data.frame to tibble to use dplyr, but it is strongly recommended. For more information about its properties please check [this](https://tibble.tidyverse.org/) out. 

```{r as_tibble}
iris_tbl <- as_tibble(iris)
```

### glimpse()

 This function is similar to str().
```{r dplyr}
# see the data frame struct
glimpse(iris_tbl)
``` 

### select()
This function allows you to select the variables you want to work with.
```{r}
# select just one variable
select(iris_tbl, Petal.Length) 
```
```{r}
# you can use commas to select one or more variables
select(iris_tbl, Petal.Length, Petal.Width)  
```
```{r}
# you can exclude variables using minus(-)
select(iris_tbl, -Species) 
```
```{r}
# with pipe
iris_tbl %>% 
  select(Petal.Length, Petal.Width)
```
```{r}
# you can use select to rearrange the variables
iris_tbl %>% 
  select(Petal.Length, Sepal.Length, 
         Petal.Width, Sepal.Width)
```

### filter()
This function returns only the rows that match some conditions. Let's see examples!
```{r filter}
# for one variable
iris_tbl %>% 
  filter(Sepal.Length > 5)
```
```{r}
# for two or more variables you can use a comma (equivalent to the AND operator)
iris_tbl %>% 
  filter(Sepal.Length > 5,  Sepal.Width < 4)
```
```{r}
# you can filter by strings with "==" operator
iris_tbl %>% 
  filter(Species == "setosa")
```

### mutate()
This function allows you to make basic operations in the variable, such as: summation, subtraction, multiplications and so on. Mutate can also create new variables in the data frame.
```{r mutate}
# you can transform a single variable
iris_tbl %>% 
  mutate(Sepal.Length = Sepal.Length * 2)

```

```{r}
# you can use commas for two or more trasformations
iris_tbl %>% 
  mutate(Sepal.Length = Sepal.Length * 2,
         Petal.Width = Petal.Width *10)
```

```{r}
# you can create any variables you want
iris_tbl %>% 
  mutate(Lengh_rate = Sepal.Length / Petal.Length,
         Width_rate = Sepal.Width / Petal.Width)
```

### working together: select(), filter(), and mutate().
```{r}
iris_tbl %>% 
  select(Petal.Length, Petal.Width) %>% 
  filter(Petal.Length > 1.5) %>% 
  mutate(Petal_rate = Petal.Length / Petal.Width)
```

### trasmute()

The transmute function also transforms or creates new variables, like mutate(). The difference betewen them is that transnmute() deletes all variables that it didn't use. In other words, transmute() is like a mutate() plus select(). Let's check it out!

```{r transmute_mutate}
# mutate() keeps all variables
iris_tbl %>% 
  mutate(Petal.Length = Petal.Length + 1)

```
```{r}
# transmute() keeps only the variables used
iris_tbl %>% 
  transmute(Petal.Length = Petal.Length + 1)
```

```{r}
# you can create new variables
iris_tbl %>% 
  transmute(Petal_rate = Petal.Length / Petal.Width)
```

```{r}
# you can trasmumte more than one variable using commas. Notice that the order of the variables in the data.frame is the same as the order written in the code.
iris_tbl %>%
  transmute(Species,
            Petal_rate = Petal.Length / Petal.Width)
```

### group_by() and summarise()

These two functions work together. The group_by() function only indexes the elements, while summarise() performs the operations. To know the all operations you can use inside summarise(), please look up `?summarise()`

```{r summarise}
# count how many observations for each Species
iris_tbl %>% 
  group_by(Species) %>% 
  summarise(Qtd = n())
```

```{r}
# mean of petal length for each species
iris_tbl %>% 
  group_by(Species) %>% 
  summarise(Petal.Lenght_mean = mean(Petal.Length))
```

## That's all folks

Have you liked it? You can share this tutorial with your colleagues using the buttons below.
If you still have any doubts or something to contribute, feel free to contact at `samuelmacedo@recife.ifpe.edu.br`.

See ya!